---
title: "BLOCK2 LAB1"
author: "Jin Yan; Donwei Ni; Collins"
date: "2022-12-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PART ONE

## QUESTION ONE
The code is given in the appendix.

Results are given below.


```{r}


# QUESTION ONE PART ONE
mis_rate1 <- c()
for(i in 1:1000){
  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(x1<x2)
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_1 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 1, nodesize = 25, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate1 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate1 <- mean(mis_rate1)
var_mis_rate1 <- var(mis_rate1)

# the mean and variable for misclassification rate when B = 1
mean_mis_rate1
var_mis_rate1


# QUESTION ONE PART TWO
mis_rate10 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(x1<x2)
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_10 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 10, nodesize = 25, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate10 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate10 <- mean(mis_rate10)
var_mis_rate10 <- var(mis_rate10)

# the mean and variable for misclassification rate when B = 10
mean_mis_rate10
var_mis_rate10





# QUESTION ONE PART THREE
mis_rate100 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(x1<x2)
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_100 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 100, nodesize = 25, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate100 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate100 <- mean(mis_rate100)
var_mis_rate100 <- var(mis_rate100)

# the mean and variable for misclassification rate when B = 10
mean_mis_rate100
var_mis_rate100
```



## QUESTION TWO
The code is given in the appendix.

Results are given below.


```{r}
# QUESTION TWO PART ONE
mis_rate1 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(x1<0.5)
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_1 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 1, nodesize = 25, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate1 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate1 <- mean(mis_rate1)
var_mis_rate1 <- var(mis_rate1)

# the mean and variable for misclassification rate when B = 1
mean_mis_rate1
var_mis_rate1


# QUESTION two PART TWO
mis_rate10 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(x1<0.5)
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_10 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 10, nodesize = 25, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate10 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate10 <- mean(mis_rate10)
var_mis_rate10 <- var(mis_rate10)
# the mean and variable for misclassification rate when B = 10
mean_mis_rate10
var_mis_rate10



# QUESTION tWO PART THREE
mis_rate100 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(x1<0.5)
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_100 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 100, nodesize = 25, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate100 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate100 <- mean(mis_rate100)
var_mis_rate100 <- var(mis_rate100)

# the mean and variable for misclassification rate when B = 10
mean_mis_rate100
var_mis_rate100


```

## QUESTION THREE
The code is given in the appendix.

Results are given below.

```{r}

# QUESTION THREE PART ONE
mis_rate1 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(((x1<0.5 & x2<0.5)|(x1>0.5 & x2>0.5)))
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_1 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 1, nodesize = 12, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate1 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate1 <- mean(mis_rate1)
var_mis_rate1 <- var(mis_rate1)

# the mean and variable for misclassification rate when B = 1
mean_mis_rate1
var_mis_rate1




# QUESTION THREE PART TWO
mis_rate10 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(((x1<0.5 & x2<0.5)|(x1>0.5 & x2>0.5)))
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_10 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 10, nodesize = 12, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate10 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate10 <- mean(mis_rate10)
var_mis_rate10 <- var(mis_rate10)
# the mean and variable for misclassification rate when B = 10
mean_mis_rate10
var_mis_rate10




# QUESTION THREE PART THREE
mis_rate100 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(((x1<0.5 & x2<0.5)|(x1>0.5 & x2>0.5)))
  trlabels<-as.factor(y)


  # QUESTION ONE
  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_100 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 100, nodesize = 12, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate100 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate100 <- mean(mis_rate100)
var_mis_rate100 <- var(mis_rate100)

# the mean and variable for misclassification rate when B = 10
mean_mis_rate100
var_mis_rate100

```

## QUESTION THREE
1.From the first two calculations, we can see the misclasification rate decreases first and then increases as the number of trees increases.

