---
title: "lab block2"
author: "Group A20"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Statement Of Contribution:**
  
Assignment 1: YAN Jin  
Assignment 2: NI Dongwei  


# PART ONE

## QUESTION ONE
The code is given in the appendix.

Results are given below.


```{r}


# QUESTION ONE PART ONE
mis_rate1 <- c()
for(i in 1:1000){
  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(x1<x2)
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_1 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 1, nodesize = 25, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate1 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate1 <- mean(mis_rate1)
var_mis_rate1 <- var(mis_rate1)

# the mean and variable for misclassification rate when B = 1
mean_mis_rate1
var_mis_rate1


# QUESTION ONE PART TWO
mis_rate10 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(x1<x2)
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_10 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 10, nodesize = 25, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate10 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate10 <- mean(mis_rate10)
var_mis_rate10 <- var(mis_rate10)

# the mean and variable for misclassification rate when B = 10
mean_mis_rate10
var_mis_rate10





# QUESTION ONE PART THREE
mis_rate100 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(x1<x2)
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_100 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 100, nodesize = 25, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate100 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate100 <- mean(mis_rate100)
var_mis_rate100 <- var(mis_rate100)

# the mean and variable for misclassification rate when B = 10
mean_mis_rate100
var_mis_rate100
```



## QUESTION TWO
The code is given in the appendix.

Results are given below.


```{r}
# QUESTION TWO PART ONE
mis_rate1 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(x1<0.5)
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_1 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 1, nodesize = 25, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate1 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate1 <- mean(mis_rate1)
var_mis_rate1 <- var(mis_rate1)

# the mean and variable for misclassification rate when B = 1
mean_mis_rate1
var_mis_rate1


# QUESTION two PART TWO
mis_rate10 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(x1<0.5)
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_10 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 10, nodesize = 25, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate10 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate10 <- mean(mis_rate10)
var_mis_rate10 <- var(mis_rate10)
# the mean and variable for misclassification rate when B = 10
mean_mis_rate10
var_mis_rate10



# QUESTION tWO PART THREE
mis_rate100 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(x1<0.5)
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_100 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 100, nodesize = 25, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate100 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate100 <- mean(mis_rate100)
var_mis_rate100 <- var(mis_rate100)

# the mean and variable for misclassification rate when B = 10
mean_mis_rate100
var_mis_rate100


```

## QUESTION THREE
The code is given in the appendix.

Results are given below.

```{r}

# QUESTION THREE PART ONE
mis_rate1 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(((x1<0.5 & x2<0.5)|(x1>0.5 & x2>0.5)))
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_1 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 1, nodesize = 12, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate1 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate1 <- mean(mis_rate1)
var_mis_rate1 <- var(mis_rate1)

# the mean and variable for misclassification rate when B = 1
mean_mis_rate1
var_mis_rate1




# QUESTION THREE PART TWO
mis_rate10 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(((x1<0.5 & x2<0.5)|(x1>0.5 & x2>0.5)))
  trlabels<-as.factor(y)



  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_10 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 10, nodesize = 12, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate10 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate10 <- mean(mis_rate10)
var_mis_rate10 <- var(mis_rate10)
# the mean and variable for misclassification rate when B = 10
mean_mis_rate10
var_mis_rate10




# QUESTION THREE PART THREE
mis_rate100 <- c()
for(i in 1:1000){

  x1<-runif(100)
  x2<-runif(100)
  trdata<-cbind(x1,x2)
  y<-as.numeric(((x1<0.5 & x2<0.5)|(x1>0.5 & x2>0.5)))
  trlabels<-as.factor(y)


  # QUESTION ONE
  train_data <- as.data.frame(trdata)
  train_data <- dplyr::mutate(train_data,trlabels)
  Forest_100 <- randomForest::randomForest(trlabels~.,data = train_data, ntree = 100, nodesize = 12, keep.forest = TRUE)
  pre <- predict(Forest_1,newdata = train_data)
  misclassification_rate <- mean(pre!=trlabels)
  mis_rate100 <- c(mis_rate1,misclassification_rate)
}

mean_mis_rate100 <- mean(mis_rate100)
var_mis_rate100 <- var(mis_rate100)

# the mean and variable for misclassification rate when B = 10
mean_mis_rate100
var_mis_rate100

```

## QUESTION THREE
1.From the first two calculations, we can see the misclasification rate decreases first and then increases as the number of trees increases.




## 2. Mixture Models

For implementing the EM algorithm for Bernoulli mixture model
First we generate a $\pi$ and $\mu$ of a Bernoulli mixture model

```{r setting, echo=FALSE}
set.seed(1234567890)
max_it <- 100 # max number of EM iterations
min_change <- 0.1 # min change in log lik between two consecutive iterations
n=1000 # number of training points
D=10 # number of dimensions
x <- matrix(nrow=n, ncol=D) # training data
true_pi <- vector(length = 3) # true mixing coefficients ; p(y)
true_mu <- matrix(nrow=3, ncol=D) # true conditional distributions ; p(xcol | y = Row)
true_pi=c(1/3, 1/3, 1/3)
true_mu[1,]=c(0.5,0.6,0.4,0.7,0.3,0.8,0.2,0.9,0.1,1)
true_mu[2,]=c(0.5,0.4,0.6,0.3,0.7,0.2,0.8,0.1,0.9,0)
true_mu[3,]=c(0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5)
plot(true_mu[1,], type="o", col="blue", ylim=c(0,1))
points(true_mu[2,], type="o", col="red")
points(true_mu[3,], type="o", col="green")
# Producing the training data x
for(i in 1:n) {
  m <- sample(1:3,1,prob=true_pi)
  for(d in 1:D) {
    x[i,d] <- rbinom(1,1,true_mu[m,d])
  } }
```

And the plot above shows the 3 $\mu$s we are using to generate data set x, with a same $\pi = \frac13$ for each $\mu$, and of course with a $Dimension = 10$.


```{r M2, echo=FALSE}
set.seed(1234567890)
M <- 2 # number of clusters
w <- matrix(nrow=n, ncol=M) # weights ; p(y = m | xi,thetaHat)
pi <- vector(length = M) # mixing coefficients
mu <- matrix(nrow=M, ncol=D) # conditional distributions
llik <- vector(length = max_it) # log likelihood of the EM iterations
# Random initialization of the parameters
pi <- runif(M,0.49,0.51)
pi <- pi / sum(pi)
for(m in 1:M) {
  mu[m,] <- runif(D,0.49,0.51)
}
pi
mu
```

First we start with cluster M = 2. And we generate a $\pi$ and $\mu$ under M = 2 as the start point for our model.

```{r M2 loop, echo=FALSE, out.width = "50%", out.height = "50%"}
set.seed(1234567890)
iterLog <- vector(length = max_it)
for(it in 1:max_it) {
  plot(mu[1,], type="o", col="blue", ylim=c(0,1))
  points(mu[2,], type="o", col="red")
  # points(mu[3,], type="o", col="green")
  #points(mu[4,], type="o", col="yellow")
  Sys.sleep(0.5)
  # E-step: Computation of the weights
  
  for (i in 1:n) {
    # px <- 0
    pxi <- 0
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
        # print(paste(i,m,d,bernXMum))
      }
      pxi <- pxi + pi[m] * bernXMum
      # print(paste(i,m,d,pxi))
    }
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
        # print(paste(i,m,d,bernXMum))
      }
      w[i,m] <- (bernXMum * pi[m]) / pxi
    }
    w[i, ] <- w[i,] /sum(w[i,])
  }
  
  
  # Your code here
  #Log likelihood computation.
  llik[it] <- 0
  for (i in 1:n) {
    pxi <- 0
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
      }
      pxi <- pxi + pi[m] * bernXMum
    }
    llik[it] <- llik[it] + log(pxi)
  }
  
  
  # Your code here
  iterLog[it] <- paste("iteration: ", it, "log likelihood: ", llik[it])
  flush.console()
  # Stop if the lok likelihood has not changed significantly
  stopFlag <- it > 1 && (llik[it] - llik[it - 1]) < min_change
  if(stopFlag) break
  #M-step: ML parameter estimation from the data and weights
  # pi mu
  pi <- apply(w, 2, mean)
  mu <- t(w) %*% x / colSums(w)
  # Your code here
}
print(iterLog[1:it])
pi
mu
plot(llik[1:it], type="o")

```

From the plots above. We can see that under this situation, $\mu$ which is equal to 0.5 every dimension seems did not being reflect. It might be that the other 2 $\mu$s are symmetrical to the 0.5 axis.  
And for every iteration the predicted $\mu$ is closer and closer to the $true \ \mu$. The rapidly raise of log-likelihood during iteration 5-7 is also reflect by the plot from significant changing of $\mu$ in corresponding iterations.
The final predicted  $\mu$ shows above is relatively close to the true value within around 10% differences.


```{r M3, echo=FALSE}
set.seed(1234567890)
M <- 3 # number of clusters
w <- matrix(nrow=n, ncol=M) # weights ; p(y = m | xi,thetaHat)
pi <- vector(length = M) # mixing coefficients
mu <- matrix(nrow=M, ncol=D) # conditional distributions
llik <- vector(length = max_it) # log likelihood of the EM iterations
# Random initialization of the parameters
pi <- runif(M,0.49,0.51)
pi <- pi / sum(pi)
for(m in 1:M) {
  mu[m,] <- runif(D,0.49,0.51)
}
pi
mu
```

Then we goes to M = 3, again generate a $\pi$ and $\mu$ under M = 3 as the start point for our model.

```{r M3 loop, echo=FALSE, out.width = "25%", out.height = "25%"}
iterLog <- vector(length = max_it)
for(it in 1:max_it) {
    # plotChoose <- c(1, 5, 6, 12, 24)
  # if (it == any(plotChoose)) {
    plot(mu[1,], type="o", col="blue", ylim=c(0,1))
    points(mu[2,], type="o", col="red")
    points(mu[3,], type="o", col="green")
    # points(mu[4,], type="o", col="yellow")
  # }
  Sys.sleep(0.5)
  # E-step: Computation of the weights
  
  for (i in 1:n) {
    # px <- 0
    pxi <- 0
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
        # print(paste(i,m,d,bernXMum))
      }
      pxi <- pxi + pi[m] * bernXMum
      # print(paste(i,m,d,pxi))
    }
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
        # print(paste(i,m,d,bernXMum))
      }
      w[i,m] <- (bernXMum * pi[m]) / pxi
    }
    w[i, ] <- w[i,] /sum(w[i,])
  }
  
  
  # Your code here
  #Log likelihood computation.
  llik[it] <- 0
  for (i in 1:n) {
    pxi <- 0
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
      }
      pxi <- pxi + pi[m] * bernXMum
    }
    llik[it] <- llik[it] + log(pxi)
  }
  
  
  # Your code here
  iterLog[it] <- paste("iteration: ", it, "log likelihood: ", llik[it])
  flush.console()
  # Stop if the lok likelihood has not changed significantly
  stopFlag <- it > 1 && (llik[it] - llik[it - 1]) < min_change
  if(stopFlag) break
  #M-step: ML parameter estimation from the data and weights
  # pi mu
  pi <- apply(w, 2, mean)
  mu <- t(w) %*% x / colSums(w)
  # Your code here
}

```

```{r M3 print, echo=FALSE, out.height="50%", out.width="50%"}
print(iterLog[1:it])
pi
mu
plot(llik[1:it], type="o")

```

We can see that the log-likelihood rises rapidly during iteration 4-7, its also clearly shown by $\mu$ values plot. Then from final result we can see that it predict the true values well, especially in comparison with following M = 4.


```{r M4, echo=FALSE}
set.seed(1234567890)
M <- 4 # number of clusters
w <- matrix(nrow=n, ncol=M) # weights ; p(y = m | xi,thetaHat)
pi <- vector(length = M) # mixing coefficients
mu <- matrix(nrow=M, ncol=D) # conditional distributions
llik <- vector(length = max_it) # log likelihood of the EM iterations
# Random initialization of the parameters
pi <- runif(M,0.49,0.51)
pi <- pi / sum(pi)
for(m in 1:M) {
  mu[m,] <- runif(D,0.49,0.51)
}
pi
mu
```

Then we change M to 4, again generate a $\pi$ and $\mu$ under M = 4 as the start point for our model.

```{r M4 loop, echo=FALSE, out.height="25%", out.width="25%"}
iterLog <- vector(length = max_it)
for(it in 1:max_it) {
  # plotChoose <- c(1, 15, 43)
  # if (it == any(plotChoose)) {
    plot(mu[1,], type="o", col="blue", ylim=c(0,1))
    points(mu[2,], type="o", col="red")
    points(mu[3,], type="o", col="green")
    points(mu[4,], type="o", col="yellow")
  # }
  Sys.sleep(0.5)
  # E-step: Computation of the weights
  
  for (i in 1:n) {
    # px <- 0
    pxi <- 0
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
        # print(paste(i,m,d,bernXMum))
      }
      pxi <- pxi + pi[m] * bernXMum
      # print(paste(i,m,d,pxi))
    }
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
        # print(paste(i,m,d,bernXMum))
      }
      w[i,m] <- (bernXMum * pi[m]) / pxi
    }
    w[i, ] <- w[i,] /sum(w[i,])
  }
  
  
  # Your code here
  #Log likelihood computation.
  llik[it] <- 0
  for (i in 1:n) {
    pxi <- 0
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
      }
      pxi <- pxi + pi[m] * bernXMum
    }
    llik[it] <- llik[it] + log(pxi)
  }
  
  
  # Your code here
  iterLog[it] <- paste("iteration: ", it, "log likelihood: ", llik[it])
  flush.console()
  # Stop if the lok likelihood has not changed significantly
  stopFlag <- it > 1 && (llik[it] - llik[it - 1]) < min_change
  if(stopFlag) break
  #M-step: ML parameter estimation from the data and weights
  # pi mu
  pi <- apply(w, 2, mean)
  mu <- t(w) %*% x / colSums(w)
  # Your code here
}
```

```{r M4 print, echo=FALSE, out.height="50%", out.width="50%"}
print(iterLog[1:it])
pi
mu
plot(llik[1:it], type="o")

```


We can see that after the rapid rising of log-likelihood, the iteration 8-15 seems still have a reasonable $\mu$ value, but then it falls into overfitting and the result $\mu$ and $\pi$ values seems not so favorable.

```{r Appendix, eval=FALSE}
# Code Appendix for part 2

set.seed(1234567890)
max_it <- 100 # max number of EM iterations
min_change <- 0.1 # min change in log lik between two consecutive iterations
n=1000 # number of training points
D=10 # number of dimensions
x <- matrix(nrow=n, ncol=D) # training data
true_pi <- vector(length = 3) # true mixing coefficients ; p(y)
true_mu <- matrix(nrow=3, ncol=D) # true conditional distributions ; p(xcol | y = Row)
true_pi=c(1/3, 1/3, 1/3)
true_mu[1,]=c(0.5,0.6,0.4,0.7,0.3,0.8,0.2,0.9,0.1,1)
true_mu[2,]=c(0.5,0.4,0.6,0.3,0.7,0.2,0.8,0.1,0.9,0)
true_mu[3,]=c(0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5)
plot(true_mu[1,], type="o", col="blue", ylim=c(0,1))
points(true_mu[2,], type="o", col="red")
points(true_mu[3,], type="o", col="green")
# Producing the training data x
for(i in 1:n) {
  m <- sample(1:3,1,prob=true_pi)
  for(d in 1:D) {
    x[i,d] <- rbinom(1,1,true_mu[m,d])
  } }

set.seed(1234567890)
M <- 2 # number of clusters
w <- matrix(nrow=n, ncol=M) # weights ; p(y = m | xi,thetaHat)
pi <- vector(length = M) # mixing coefficients
mu <- matrix(nrow=M, ncol=D) # conditional distributions
llik <- vector(length = max_it) # log likelihood of the EM iterations
# Random initialization of the parameters
pi <- runif(M,0.49,0.51)
pi <- pi / sum(pi)
for(m in 1:M) {
  mu[m,] <- runif(D,0.49,0.51)
}
pi
mu

set.seed(1234567890)
iterLog <- vector(length = max_it)
for(it in 1:max_it) {
  plot(mu[1,], type="o", col="blue", ylim=c(0,1))
  points(mu[2,], type="o", col="red")
  # points(mu[3,], type="o", col="green")
  #points(mu[4,], type="o", col="yellow")
  Sys.sleep(0.5)
  # E-step: Computation of the weights
  
  for (i in 1:n) {
    # px <- 0
    pxi <- 0
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
        # print(paste(i,m,d,bernXMum))
      }
      pxi <- pxi + pi[m] * bernXMum
      # print(paste(i,m,d,pxi))
    }
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
        # print(paste(i,m,d,bernXMum))
      }
      w[i,m] <- (bernXMum * pi[m]) / pxi
    }
    w[i, ] <- w[i,] /sum(w[i,])
  }
  
  
  # Your code here
  #Log likelihood computation.
  llik[it] <- 0
  for (i in 1:n) {
    pxi <- 0
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
      }
      pxi <- pxi + pi[m] * bernXMum
    }
    llik[it] <- llik[it] + log(pxi)
  }
  
  
  # Your code here
  iterLog[it] <- paste("iteration: ", it, "log likelihood: ", llik[it])
  flush.console()
  # Stop if the lok likelihood has not changed significantly
  stopFlag <- it > 1 && (llik[it] - llik[it - 1]) < min_change
  if(stopFlag) break
  #M-step: ML parameter estimation from the data and weights
  # pi mu
  pi <- apply(w, 2, mean)
  mu <- t(w) %*% x / colSums(w)
  # Your code here
}
print(iterLog[1:it])
pi
mu
plot(llik[1:it], type="o")

set.seed(1234567890)
M <- 3 # number of clusters
w <- matrix(nrow=n, ncol=M) # weights ; p(y = m | xi,thetaHat)
pi <- vector(length = M) # mixing coefficients
mu <- matrix(nrow=M, ncol=D) # conditional distributions
llik <- vector(length = max_it) # log likelihood of the EM iterations
# Random initialization of the parameters
pi <- runif(M,0.49,0.51)
pi <- pi / sum(pi)
for(m in 1:M) {
  mu[m,] <- runif(D,0.49,0.51)
}
pi
mu


iterLog <- vector(length = max_it)
for(it in 1:max_it) {
    # plotChoose <- c(1, 5, 6, 12, 24)
  # if (it == any(plotChoose)) {
    plot(mu[1,], type="o", col="blue", ylim=c(0,1))
    points(mu[2,], type="o", col="red")
    points(mu[3,], type="o", col="green")
    # points(mu[4,], type="o", col="yellow")
  # }
  Sys.sleep(0.5)
  # E-step: Computation of the weights
  
  for (i in 1:n) {
    # px <- 0
    pxi <- 0
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
        # print(paste(i,m,d,bernXMum))
      }
      pxi <- pxi + pi[m] * bernXMum
      # print(paste(i,m,d,pxi))
    }
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
        # print(paste(i,m,d,bernXMum))
      }
      w[i,m] <- (bernXMum * pi[m]) / pxi
    }
    w[i, ] <- w[i,] /sum(w[i,])
  }
  
  
  # Your code here
  #Log likelihood computation.
  llik[it] <- 0
  for (i in 1:n) {
    pxi <- 0
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
      }
      pxi <- pxi + pi[m] * bernXMum
    }
    llik[it] <- llik[it] + log(pxi)
  }
  
  
  # Your code here
  iterLog[it] <- paste("iteration: ", it, "log likelihood: ", llik[it])
  flush.console()
  # Stop if the lok likelihood has not changed significantly
  stopFlag <- it > 1 && (llik[it] - llik[it - 1]) < min_change
  if(stopFlag) break
  #M-step: ML parameter estimation from the data and weights
  # pi mu
  pi <- apply(w, 2, mean)
  mu <- t(w) %*% x / colSums(w)
  # Your code here
}

print(iterLog[1:it])
pi
mu
plot(llik[1:it], type="o")

set.seed(1234567890)
M <- 4 # number of clusters
w <- matrix(nrow=n, ncol=M) # weights ; p(y = m | xi,thetaHat)
pi <- vector(length = M) # mixing coefficients
mu <- matrix(nrow=M, ncol=D) # conditional distributions
llik <- vector(length = max_it) # log likelihood of the EM iterations
# Random initialization of the parameters
pi <- runif(M,0.49,0.51)
pi <- pi / sum(pi)
for(m in 1:M) {
  mu[m,] <- runif(D,0.49,0.51)
}
pi
mu


iterLog <- vector(length = max_it)
for(it in 1:max_it) {
  # plotChoose <- c(1, 15, 43)
  # if (it == any(plotChoose)) {
    plot(mu[1,], type="o", col="blue", ylim=c(0,1))
    points(mu[2,], type="o", col="red")
    points(mu[3,], type="o", col="green")
    points(mu[4,], type="o", col="yellow")
  # }
  Sys.sleep(0.5)
  # E-step: Computation of the weights
  
  for (i in 1:n) {
    # px <- 0
    pxi <- 0
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
        # print(paste(i,m,d,bernXMum))
      }
      pxi <- pxi + pi[m] * bernXMum
      # print(paste(i,m,d,pxi))
    }
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
        # print(paste(i,m,d,bernXMum))
      }
      w[i,m] <- (bernXMum * pi[m]) / pxi
    }
    w[i, ] <- w[i,] /sum(w[i,])
  }
  
  
  # Your code here
  #Log likelihood computation.
  llik[it] <- 0
  for (i in 1:n) {
    pxi <- 0
    for (m in 1:M) {
      bernXMum <- 1
      for (d in 1:D) {
        bernXMum <- bernXMum * (mu[m,d]^x[i,d]) * (1-mu[m,d])^(1-x[i,d])
      }
      pxi <- pxi + pi[m] * bernXMum
    }
    llik[it] <- llik[it] + log(pxi)
  }
  
  
  # Your code here
  iterLog[it] <- paste("iteration: ", it, "log likelihood: ", llik[it])
  flush.console()
  # Stop if the lok likelihood has not changed significantly
  stopFlag <- it > 1 && (llik[it] - llik[it - 1]) < min_change
  if(stopFlag) break
  #M-step: ML parameter estimation from the data and weights
  # pi mu
  pi <- apply(w, 2, mean)
  mu <- t(w) %*% x / colSums(w)
  # Your code here
}

print(iterLog[1:it])
pi
mu
plot(llik[1:it], type="o")


```

