---
title: "Lab01 Assignment 3"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3. Logistic regression and basis function expansion
The data file **pima-indians-diabetes.csv** contains information about the onset of
diabetes within 5 years in Pima Indians given medical details. The variables are (in
the same order as in the dataset):

1. Number of times pregnant.  
2. Plasma glucose concentration a 2 hours in an oral glucose tolerance test.  
3. Diastolic blood pressure (mm Hg).  
4. Triceps skinfold thickness (mm).  
5. 2-Hour serum insulin (mu U/ml).  
6. Body mass index (weight in kg/(height in m)^2).  
7. Diabetes pedigree function.  
8. Age (years).  
9. Diabetes (0=no or 1=yes)  

1.Make a scatterplot showing a Plasma glucose concentration on Age where observations are colored by Diabetes levels. Do you think that Diabetes is easy to classify by a standard logistic regression model that uses these two variables as features? Motivate your answer.


```{r library, include=FALSE}
library(tidyverse)
library(ggplot2)
library(shipunov)
```
```{r 3.1, echo=TRUE}

data_PIdiabetes <- read.csv("pima-indians-diabetes.csv", col.names = c(1:9))

data_PIdiabetes$X9 <- factor(data_PIdiabetes$X9)

ggplot(data = data_PIdiabetes, aes(x = X8, y = X2, colour = X9)) + 
  geom_point() + 
  labs(x = "Age (years)", y = "Plasma glucose concentration",
       colour  = "Diabetes (0=no or 1=yes)" )

```


- No! Motivation: Looking at the plot, one can see that the data points are mixed for the group with Diabetes level 1 compared to those with level 0. And the decision boundary does not seems can be linear enough.
At such one can conclude that it will be hard to classify Diabetes by a standard logistic regression if we want a low error rate

2.Train a logistic regression model with $y$ = Diabetes as target $x_1$ = Plasma glucose
concentration and $x_2$ = Age as features and make a prediction for all observations
by using $r$ = 0.5 as the classification threshold. Report the probabilistic equation
of the estimated model (i.e., how the target depends on the features and the
estimated model parameters probabilistically). Compute also the training
misclassification error and make a scatter plot of the same kind as in step 1 but
showing the predicted values of Diabetes as a color instead. Comment on the
quality of the classification by using these results

```{r 3.2 split, eval=FALSE, include=FALSE}
set.seed(24601)

n <- nrow(data_PIdiabetes)

r_train <- sample(1:n, floor(n*0.5))
train <-  data_PIdiabetes[r_train, ]

r_nont <- setdiff(1:n, r_train)
r_valid <- sample(r_nont, floor(n*0.25))
valid <- data_PIdiabetes[r_valid, ]
 
test <- data_PIdiabetes[setdiff(r_nont, r_valid), ]

```

- Logistic regression analysis belongs to the class of generalized linear models.
In R generalized linear models are handled by the glm() function. The function is
written as glm(response ~ predictor, family = binomial(link = "logit"), data).
Please note that logit is the default for binomial; thus, we do not have to type it explicitly.
The glm() function returns a model object, therefore we may apply extractor functions, such as summary(),
fitted() or predict, among others, on it. However, please note that the output numbers are on the logit scale.
To actually predict probabilities we need to provide the predict() function an
additional argument type = "response"

```{r 3.2 glm}
# X2. Plasma glucose concentration a 2 hours in an oral glucose tolerance test.
# X8. Age (years).
# X9. Diabetes (0=no or 1=yes).

glm_diabete <- glm(X9 ~ X2 + X8, data_PIdiabetes, family = binomial)

fit <- ifelse(fitted(glm_diabete) >= 0.5, 1, 0)

```

```{r 3.2 probabilistic equation}
glm_diabete$coefficients
```
- we have
$$g(x) = \cfrac{1}{1 + e^{-\theta^T x}} $$
and use coefficients we got:
$$g(x) = \cfrac{1}{1 + e^{-(-5.89785793 + 0.03558250x_1 + 0.02450157x_2)}} $$
while $x_1$ = Plasma glucose concentration and $x_2$ = age.

```{r 3.2 misclassification plot}

tq1 <- table(fit, data_PIdiabetes$X9)
tq1
# confusion matrix
1 - (sum(diag(tq1)) / sum(tq1))
# misclassification error


data_PIdiabetes_q1 <- data_PIdiabetes
data_PIdiabetes_q1$X10 <- factor(fit)

ggplot(data = data_PIdiabetes_q1, aes(x = X8, y = X2, colour = X10)) + 
  geom_point() + 
  labs(x = "Age (years)", y = "Plasma glucose concentration",
       colour = "Predicted Diabetes (0=no or 1=yes)" )


```

- From the glm model, Both Plasma glucose and Age have a significant influence on Diabetes. From the results of this classification, we can see that our misclassification error is about 26.6%. This is not good enough especially for diabetes classification. We don't want those who have diabetes be wrongly classified as being healthy.


3.Use the model estimated in step 2 to a) report the equation of the decision
boundary between the two classes b) add a curve showing this boundary to the
scatter plot in step 2. Comment whether the decision boundary seems to catch the
data distribution well.

- when r = 0.5, we have
$$\cfrac{1}{1 + e^{-\theta^T x}} = 0.5$$
then we have $\theta^T x = 0$  , i.e. $\theta_0 + \theta_1x_1 + \theta_2x_2 =0$  ($x_1: concentration, x_2: age$)  
thus $slope = \cfrac{-\theta_2}{\theta_1}$, $intercept = \cfrac{-\theta_0}{\theta_1}$


```{r 3.3}
theta <- coef(glm_diabete)

slope <- (-theta[3]) / theta[2]
intercept <- (-theta[1] ) / (theta[2]) 

ggplot(data = data_PIdiabetes_q1, aes(x = X8, y = X2, colour = X10)) + 
  geom_point() + 
  geom_abline(slope = slope, intercept = intercept) +
  
  labs(x = "Age (years)", y = "Plasma glucose concentration",
       colour  = "Predicted Diabetes (0=no or 1=yes)")


```

- Comments:I think this decision boundary does well for separating the data belonging two classes, even though not all data are put into the right sides of the boundary line.



4.Make same kind of plots as in step 2 but use thresholds r = 0.2 and r = 0.8. By
using these plots, comment on what happens with the prediction when r value
changes.  


```{r 3.4 r0.2}
fit_r02 <- ifelse(fitted(glm_diabete) >= 0.2, 1, 0)

tr02 <- table(fit_r02, data_PIdiabetes$X9)
tr02
#confusion matrix
1 - (sum(diag(tr02)) / sum(tr02))
#misclassification error


data_PIdiabetes_r02 <- data_PIdiabetes
data_PIdiabetes_r02$X10 <- factor(fit_r02)

slope <- (-theta[3]) / theta[2]
intercept <- (-log(4) - theta[1] ) / (theta[2]) 

ggplot(data = data_PIdiabetes_r02, aes(x = X8, y = X2, colour = X10)) + 
  geom_point() + 
  geom_abline(slope = slope, intercept = intercept) +
  
  labs(x = "Age (years)", y = "Plasma glucose concentration",
       colour  = "Predicted Diabetes (0=no or 1=yes)", title = "r = 0.2" )

```

```{r 3.4 r0.8}

fit_r08 <- ifelse(fitted(glm_diabete) >= 0.8, 1, 0)


tr08 <- table(fit_r08, data_PIdiabetes$X9)
tr08
#confusion matrix
1 - (sum(diag(tr08)) / sum(tr08))
#misclassification error

data_PIdiabetes_r08 <- data_PIdiabetes
data_PIdiabetes_r08$X10 <- factor(fit_r08)

slope <- (-theta[3]) / theta[2]
intercept <- (log(4) - theta[1] ) / (theta[2]) 

ggplot(data = data_PIdiabetes_r08, aes(x = X8, y = X2, colour = X10)) + 
  geom_point() + 
  geom_abline(slope = slope, intercept = intercept) +
  
  labs(x = "Age (years)", y = "Plasma glucose concentration",
       colour  = "Predicted Diabetes (0=no or 1=yes)", title = "r = 0.8" )


```

- Comment: When r is 0.8, we can find that the method is more likely to regard the patients with diabetes as healthy. The probability is 87%. By contrast, when we use r = 0.2 as the threshold, we can find the patients with more probability, which is 91%.



```{r}
242/(242+25)
231/(231+36)

```

5.Perform a basis function expansion trick by computing new features $z_1 = x_1^4, z_2 = x_1^3x_2^1, z_3 = x_1^2x_2^2, z_4 = x_1^1x_2^3, z_5 = x_2^4$, adding them to the data set and
then computing a logistic regression model with y as target and
$x_1, x_2, z_1, …,z_5$ as features. Create a scatterplot of the same kind as in step 2
for this model and compute the training misclassification rate. What can you
say about the quality of this model compared to the previous logistic
regression model? How have the basis expansion trick affected the shape of
the decision boundary and the prediction accuracy?  


```{r 3.5}
data_PIdiabetes_q5 <-  data_PIdiabetes %>%  select(c(X2,X8,X9))
# X2,X8,X9
# 2. Plasma glucose concentration a 2 hours in an oral glucose tolerance test.
# 8. Age (years).
# 9. Diabetes (0=no or 1=yes).

data_PIdiabetes_q5$Z1 <- data_PIdiabetes_q5$X2^4
data_PIdiabetes_q5$Z2 <- data_PIdiabetes_q5$X2^3 * data_PIdiabetes_q5$X8^1
data_PIdiabetes_q5$Z3 <- data_PIdiabetes_q5$X2^2 * data_PIdiabetes_q5$X8^2
data_PIdiabetes_q5$Z4 <- data_PIdiabetes_q5$X2^1 * data_PIdiabetes_q5$X8^3
data_PIdiabetes_q5$Z5 <- data_PIdiabetes_q5$X8^4

glm_diabete_q5 <- glm(X9 ~ X2 + X8 + Z1 + Z2 + Z3 + Z4 + Z5, data_PIdiabetes_q5, family = binomial)

fit <- ifelse(fitted(glm_diabete_q5) >= 0.5, 1, 0)

tq5 <- table(fit, data_PIdiabetes_q5$X9)
tq5
# confusion matrix
1 - (sum(diag(tq5)) / sum(tq5))
# misclassification error

data_PIdiabetes_q5$X10 <- factor(fit)

ggplot(data = data_PIdiabetes_q5, aes(x = X8, y = X2, colour = X10)) +
  geom_point() +
  labs(x = "Age", y = "Plasma glucose concentration",
       colour  = "Predicted Diabetes (0=no or 1=yes)")

```

- Comment: When we increase the number of inputs and r equals 0.5, the eventual value of misclassification rate is similar to the one in the second question. The boundary line is no longer a straight line but a curve.  
Even though this model have the lowest error rate of all models we have, we still don't consider this one as the best model we have. Because the most important thing is that we can accept that healthy be misclassified as with diabetes, but not the other way round. Thus the r = 0.2 model is the most practical one.
